---
title: "CS512_project1"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r}
defaultW <- getOption("warn") 

options(warn = -1) 
```


<h1>Import Packages</h1>
```{r}
library(factoextra)
library(ClusterR)
library(Hmisc)
library(ggplot2)
library(reshape2)
library(mclust)
library(cluster)
```

<h1>Load Data</h1>
```{r}
data <- read.csv('data/customers.csv')
```


<h1>Exploratory Analysis</h1>
<h2>Examine Statistics</h2>
```{r}
summary(data)
describe(data)
```
<h2>Examine First Few Samples</h2>
```{r}
head(data)
```
<h2>Compute and Examine Correlation Coefficients</h2>
```{r}
corr_mat <- cor(data)
corr_mat
```
<h1>Data Visualization</h1>
<h2>Heat Map of Correlation Matrix</h2>
```{r}
melted_corr_mat <- melt(corr_mat)
ggplot(data = melted_corr_mat, aes(x=Var1, y=Var2, fill=value)) + geom_tile()
```
<h2>Perform PCA and Visualize Bi-Plot</h2>
```{r}
#perform PCA
reduced_data <- princomp(data)

#create biplot to visualize results of PCA
biplot(reduced_data)
```
<h1>Data Preprocessing</h1>
<h2>Remove Categorical Variables</h2>
```{r}
keep_cols <- c("Fresh", "Milk", "Grocery", "Frozen", "Detergents_Paper","Delicatessen")
good_data <- data[keep_cols]
head(good_data)
```
<h2>Remove Highly Correlated Variable</h2>
```{r}
keep_cols <- c("Fresh", "Milk", "Frozen", "Detergents_Paper","Delicatessen")
good_data <- good_data[keep_cols]
head(good_data)
```
<h1>Clustering Algorithms</h1>
<h2>K-means Clustering</h2>
<h3>Building Models</h3>
```{r}
# No. of Clusters = 3
start_time3 <- Sys.time()
cust_kmeans3 = kmeans(good_data, 3)
end_time3 <- Sys.time()

# No. of Clusters = 20
start_time20 <- Sys.time()
cust_kmeans20 = kmeans(good_data, 20)
end_time20 <- Sys.time()

# No. of Clusters = 50
start_time50 <- Sys.time()
cust_kmeans50 = kmeans(good_data, 50)
end_time50 <- Sys.time()
```

<h3>Visualize Results</h3>
```{r}
# No. of Clusters = 3
fviz_cluster(cust_kmeans3, good_data, choose.vars = TRUE, stand = FALSE, show.clust.cent = TRUE)

# No. of Clusters = 20
fviz_cluster(cust_kmeans20, good_data, choose.vars = TRUE, stand = FALSE, show.clust.cent = TRUE)

# No. of Clusters = 50
fviz_cluster(cust_kmeans50, good_data, choose.vars = TRUE, stand = FALSE, show.clust.cent = TRUE)
```

<h3>Compute Metrics</h3>
<h4>Silhouette Coefficient</h4>
```{r}
# No. of Clusters = 3
Kmeans_sil3 = mean(silhouette(cust_kmeans3$cluster, dist(good_data))[,3])
print(Kmeans_sil3)

# No. of Clusters = 20
Kmeans_sil20 = mean(silhouette(cust_kmeans20$cluster, dist(good_data))[,3])
print(Kmeans_sil20)

# No. of Clusters = 50
Kmeans_sil50 = mean(silhouette(cust_kmeans50$cluster, dist(good_data))[,3])
print(Kmeans_sil50)
```
<h4>Execution Time</h4>
```{r}
# No. of Clusters = 3
kmeans3_exe_time <- end_time3 - start_time3
print(kmeans3_exe_time)

# No. of Clusters = 20
kmeans20_exe_time <- end_time20 - start_time20
print(kmeans20_exe_time)

# No. of Clusters = 50
kmeans50_exe_time <- end_time50 - start_time50
print(kmeans50_exe_time)
```

<h2>Hierarchical Clustering (DIANA)</h2>
<h3>Building Models</h3>
```{r}
# No. of Clusters = 3
start_time3 <- Sys.time()
cust_tree3 = hcut(good_data, k = 3, isdiss = FALSE, hc_func = "diana", hc_metric = "euclidean", stand = FALSE, graph = TRUE)
end_time3 <- Sys.time()

# No. of Clusters = 20
start_time20 <- Sys.time()
cust_tree20 = hcut(good_data, k = 20, isdiss = FALSE, hc_func = "diana", hc_metric = "euclidean", stand = FALSE, graph = TRUE)
end_time20 <- Sys.time()

# No. of Clusters = 50
start_time50 <- Sys.time()
cust_tree50 = hcut(good_data, k = 50, isdiss = FALSE, hc_func = "diana", hc_metric = "euclidean", stand = FALSE, graph = TRUE)
end_time50 <- Sys.time()
```
<h3>Visualize Results</h3>
```{r}
# No. of Clusters = 3
fviz_cluster(cust_tree3)

# No. of Clusters = 20
fviz_cluster(cust_tree20)

# No. of Clusters = 50
fviz_cluster(cust_tree50)
```

<h3>Compute Metrics</h3>
<h4>Silhouette Coefficient</h4>
```{r}
# No. of Clusters = 3
diana3_sil = mean(silhouette(cust_tree3$cluster, dist(good_data))[,3])
print(diana3_sil)

# No. of Clusters = 20
diana20_sil = mean(silhouette(cust_tree20$cluster, dist(good_data))[,3])
print(diana20_sil)

# No. of Clusters = 50
diana50_sil = mean(silhouette(cust_tree50$cluster, dist(good_data))[,3])
print(diana50_sil)
```
<h4>Execution Time</h4>
```{r}
# No. of Clusters = 3
diana3_exe_time <- end_time3 - start_time3
print(diana3_exe_time)

# No. of Clusters = 20
diana20_exe_time <- end_time20 - start_time20
print(diana20_exe_time)

# No. of Clusters = 50
diana50_exe_time <- end_time50 - start_time50
print(diana50_exe_time)
```

<h2>Gaussian Mixture Model Clustering</h2>
<h3>Building Models</h3>
```{r}
# No. of Clusters = 3
start_time3 <- Sys.time()
gmm3 <- Mclust(good_data, G = 3)
end_time3 <- Sys.time()

# No. of Clusters = 20
start_time20 <- Sys.time()
gmm20 <- Mclust(good_data, G = 20)
end_time20 <- Sys.time()

# No. of Clusters = 50
start_time20 <- Sys.time()
gmm50 <- Mclust(good_data, G = 50)
end_time20 <- Sys.time()
```

<h3>Visualize Results</h3>
```{r}
# No. of Clusters = 3
fviz_cluster(gmm3)

# No. of Clusters = 20
fviz_cluster(gmm20)

# No. of Clusters = 50
fviz_cluster(gmm50)
```

<h3>Compute Metrics</h3>
<h4>Silhouette Coefficient</h4>
```{r}
# No. of Clusters = 3
gmm3_sil <- mean(silhouette(gmm3$classification, dist(good_data))[,3])
print(gmm3_sil)

# No. of Clusters = 20
gmm20_sil <- mean(silhouette(gmm20$classification, dist(good_data))[,3])
print(gmm20_sil)

# No. of Clusters = 50
gmm50_sil <- mean(silhouette(gmm50$classification, dist(good_data))[,3])
print(gmm50_sil)
```
<h4>Execution Time</h4>
```{r}
# No. of Clusters = 3
gmm3_exe_time <- end_time3 - start_time3
print(gmm3_exe_time)

# No. of Clusters = 20
gmm20_exe_time <- end_time20 - start_time20
print(gmm20_exe_time)

# No. of Clusters = 50
gmm50_exe_time <- end_time50 - start_time50
print(gmm50_exe_time)
```

<h1>Comparison</h1>
<h2>Comparative Analysis of Silhouette Coefficients</h2>
<style>
	.demo {
		border:1px solid #C0C0C0;
		border-collapse:collapse;
		padding:5px;
	}
	.demo th {
		border:1px solid #C0C0C0;
		padding:5px;
		background:#F0F0F0;
	}
	.demo td {
		border:1px solid #C0C0C0;
		padding:5px;
	}
</style>
<table class="demo">
	<caption><b>Silhouette Coefficients</b></caption>
	<thead>
	<tr>
		<th><br></th>
		<th>K-means</th>
		<th>DIANA</th>
		<th>GMM</th>
	</tr>
	</thead>
	<tbody>
	<tr>
		<td><b>3 Clusters</b>&nbsp;</td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">0.3537846</pre></td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;"><b>0.6269519</b></pre></td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">0.1048809</pre></td>
	</tr>
	<tr>
		<td><b>20 Clusters</b></td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">0.2924626</pre></td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">0.3896082</pre></td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">-0.03335207</pre></td>
	</tr>
	<tr>
		<td><b>50 Clusters</b></td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">0.2334977</pre></td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">0.242066</pre></td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">0.07322262</pre></td>
	</tr>
	</tbody>
</table>

<h2>Comparative Analysis of Execution Time</h2>
<style>
	.demo {
		border:1px solid #C0C0C0;
		border-collapse:collapse;
		padding:5px;
	}
	.demo th {
		border:1px solid #C0C0C0;
		padding:5px;
		background:#F0F0F0;
	}
	.demo td {
		border:1px solid #C0C0C0;
		padding:5px;
	}
</style>
<table class="demo">
	<caption><b>Execution Time (in seconds)</b></caption>
	<thead>
	<tr>
		<th><br></th>
		<th>K-means</th>
		<th>DIANA</th>
		<th>GMM</th>
	</tr>
	</thead>
	<tbody>
	<tr>
		<td><b>3 Clusters</b>&nbsp;</td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">0.003316879</pre></td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">10.12301</pre></td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">0.2640531</pre></td>
	</tr>
	<tr>
		<td><b>20 Clusters</b></td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">0.002763987</pre></td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">10.41367</pre></td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">0.8423929</pre></td>
	</tr>
	<tr>
		<td><b>50 Clusters</b></td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">0.005206108</pre></td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">10.77101</pre></td>
		<td><pre data-ordinal="1" style="font-family: Consolas, &quot;Lucida Console&quot;, monospace, monospace; font-size: 13.8667px; line-height: 1.1; margin-bottom: 0px; color: rgb(0, 0, 0); margin-top: 0px; white-space: pre-wrap;">15.20578</pre></td>
	</tr>
	</tbody>
</table>

```{r}
options(warn = defaultW)
```

